### 2020-12-10 Thesis-Chapter3-PolicyMaking.tex
Editorial improvements

Before going into detail about the arguments used in the \ac{EA} debate, it is worth studying \acp{wicked-prob} in
greater detail. What are they, and why is the case of encryption, privacy, and \ac{EA} in this category? What approaches
to tackling \acp{wicked-prob} succeed and fail? How can we collectively situate ourselves to make real progress? This
chapter seeks to answer these questions.

@@

\Acp{wicked-prob} were previously introduced as pernicious and tricky issues that resist straightforward solutions.
In this section we analyze the nature of wicked problems and strategies for tackling them.

@@

Rittel's categorization of \acp{wicked-prob} grew out of frustration with the problems' resistance to classical problem
solving methods. Since the Enlightenment, society has applied the scientific method to problems of every kind; the
sweeping application of scientific analysis has delivered reliable clean water, improved crop yields, shaped government
structures, and bestowed material wealth previously unimaginable. With our material problems largely solved in the
twentieth century, believers in the power of reason thought this progress would march on in the realm of public
planning. Policymaking would function by setting goals, identifying problems, evaluating alternatives, implementing
solutions, and analyzing outcomes in order to correct errors. Functioning as a continuous process, this approach was
primed to revolutionize governing the same way it did industry, agriculture, and economics---until it didn't. In the
context of what he describes as an ``anti-professional movement,'' Rittel explains how the scientific method has failed:

@@

When applied to social problems, the prescribed method---here defined as setting goals, identifying problems, evaluating
alternatives, implementing solutions, and analyzing outcomes---fails at every step. In the U.S., a nation with countless
cultures and subcultures which are effectively represented by two mutually hostile political parties, agreeing on goals
is a challenge in itself. When goals are set, we often discover that we are contending with \acp{wicked-prob} that defy
the process at each remaining step. Rittel provides a list of ten characteristics of problems in this category
\cite{rittel_dilemmas_1973}:

@@

Based on this description, \acp{wicked-prob} are impossible to definitively identify, their alternative solutions are
infinite in number but cannot be proved or independently tested, solution implementations carry large risks, and outcome
analysis is subject to interpretation. The policy version of the scientific method cannot be used under these
conditions. The ``wicked'' problem's counterpart is the ``tame'' problem. ``Tame'' does not imply easy; it only means
that the scientific approach will be effective.

Several of \acp{wicked-prob}' characteristics are results of the fact that we have no accurate predictive model of the
world and human behavior at large. This does not need to be expounded upon.

The more important insight is one step removed: precisely because there is an inexhaustible set of potential solutions,
the problem definition---that is, the set of information required to produce a solution---is not self-contained. Each
proposed solution demands research and context to be fed back into the problem definition. This feedback from proposal
to definition violates the linearity of the traditional approach. One cannot reason from problem statement to proposals
to a solution; instead, one is forced to constantly refine the problem statement based on the content of proposals
themselves.

A 2018 study by the Australian government uses climate change as an example of a \ac{wicked-prob}
\cite{commission_tackling_2018}. Say the problem is initially formulated as long-term change to the environment caused
by the effects of accumulating greenhouse gasses. Responses typically take one of three broad forms: profligate behavior
in consumeristic societies must be reigned in at a local and personal level; global inter-governmental coordination is
the only solution, as individual changes will have no impact; or the situation is overblown by idealists and
power-mongers, and technological progress and adaptive markets will handle any negative effects that come to pass
\cite{commission_tackling_2018}. Which response is correct? One cannot tell from the problem statement. How much do
individual choices contribute to greenhouse gas emissions? How effective are international accords? How costly will the
changes be, and how capable is technology to respond? Evaluating the validity of each proposal requires updates to the
problem statement itself.

@@

\Acp{wicked-prob} always consist of subproblems, which may be tame or wicked in and of themselves. For example, climate
change has the tame subproblem of developing clean energy technology, and the wicked subproblem of implementing any
solution in a way that doesn't leave vast swaths of people behind. \Acp{wicked-prob} quickly grow in complexity with the
number of subproblems that comprise them. Truly imposing \acp{wicked-prob} are a tangle of contributing factors, and
despite the presence of tame elements, they pose a large number of difficult challenges.

@@

\begin{enumerate}
  \item There is no formulation that encapsulates the problem of encryption's interplay with privacy, security,
        safety, and trust.
  \item Balancing each value in the face of constantly evolving technology is never ending.
  \item \ac{EA} or other proposals cannot definitively solve the problem.
  \item \ac{EA} or other proposals cannot be objectively tested.
  \item Every policy implementation has irreversible effects.
  \item There is an inexhaustible set of potential solutions to the problem.
  \item Solutions from other domains do not apply directly to this problem.
  \item The need for \ac{EA} or some alternative is a symptom of differing values, rapid technological change, and
        criminal behavior.
  \item The problem can be framed as insufficient investigatory access to digital data (presupposing \ac{EA} as the
        solution), outdated cyberlaw (presupposing legal hacking or compelled password disclosure as solutions), and
        more.
  \item The decisions of regulators and technologists alike have real impacts in the world today.
\end{enumerate}

Cybersecurity law scholar Alan Rozenshtein argues for treating \ac{encryption} and \ac{EA} as a \ac{wicked-prob} at
length \cite{rozenshtein_wicked_2018}. He breaks the root issues down into three categories. First, that there is
disagreement on goals (or even basic premises). Sides can't agree whether things are ``going dark'' or we're in the
``golden age for surveillance'' or how much to value competing notions of security. Second, that information is
``uncertain and diffuse.'' There are insufficient statistics on encryption's effect on investigations, and though the
consensus is that we are not presently capable of acceptably secure \ac{EA}, that is not certain to always be the case,
especially as the area is under-researched. Third, that the problem cannot be definitively solved. Evolving values and
technology mean that the subject is always up for renegotiation.

\Ac{encryption} technology is particularly sensitive to the property that policy implementation has irreversible
effects. Technology deployments have long tails, and attackers have the ability to record and store data for later
analysis. In one investigation, Australian authorities cracked a cold case based on evidence acquired from a mobile
device that they cracked five years after it was seized \cite{evans_arrests_2020}. Lawful authorities wouldn't be the
only ones to benefit from a policy miscalculation could negatively impact security for a very long time.

@@

Embracing reality is the first step to dealing with it \cite{baker_2019}. We have by now embraced the reality of
\acp{wicked-prob}. In the next two sections, we look at strategies for dealing with it.

@@

The \ac{classical-method} \cite{feeley_judicial_2000} (or ``the modern-classical model of planning''
\cite{rittel_dilemmas_1973}, ``the rational-comprehensive method'' \cite{lindblom_muddling_1959}, ``traditional policy
analysis'' \cite{rozenshtein_wicked_2018}, or ``linear thinking'' \cite{commission_tackling_2018}) has already been
introduced. It is the reason-based method that functions by setting goals, identifying problems, evaluating
alternatives, implementing solutions, and analyzing outcomes in order to correct errors. \myfig{fig-classical-method}
illustrates the approach in the context of encryption and \ac{EA}. It has a purely linear flow from problem to solution
except for the ``refinement'' step, in which the method analyzes outcomes and corrects errors. However, it is important
to note that refinement is always in-kind---it represents a reinforcement, as opposed to a reassessment, of the chosen
solution.

@@

The shortcomings of the \ac{classical-method} were discussed in \mysec{wicked-characteristics}. It fails due to
disagreement over goals, the dependence of the problem definition on the alternatives generated (violating the linear
flow), the inability to evaluate alternatives, and the lack of a definitive stopping rule.

@@

Lindblom's ``\ac{muddling-through}'' operates by taking successive steps chosen through comparative analysis. The
alternatives selected for comparison are defined relative to the status quo and are close enough to one another to be
able to be analyzed on the margin. This is done first out of practical necessity, due to our inability to rationally
predict policy outcomes, and second out of political realism, as non-incremental changes are usually politically
impossible to impose in a democratic system \cite{lindblom_muddling_1959}. The formality of the process varies;
policymakers may use this method consciously, with considerable comparative analysis, or unconsciously, led by
intuition. \myfig{fig-muddling-through} illustrates this method.

@@

Despite its strengths, \ac{incrementalism} has its weaknesses at addressing \acp{wicked-prob}. Its main weakness is its
lack of high level strategic analysis. It is incapable of drastic change, which is sometimes necessary. By Lindblom's
admission, it lacks a safeguard for consideration of all relevant values and may ``overlook excellent policies for no
other reason than that they are not suggested by the chain of successive policy steps leading up to the present''
\cite{lindblom_muddling_1959}. Analysis based only on one present ``status quo'' to the next can result in messy policy
that none are particularly fond of---``As Lindblom's sobriquet suggests, it often [leads] to a considerable muddle''
\cite{feeley_judicial_2000}. Consider application of the \acl{CFAA} as one such muddle \cite{wolff_computer_2016}.

@@

The \ac{classical-method} and \ac{incrementalism} are not the only styles of policymaking. They represent perhaps two
extremes on a spectrum of rational planning and intuitive muddling. Both have their strengths and weaknesses. One may
intuit that a reasonable strategy is the selective use of both approaches according to the situation, and indeed this
has also been formally suggested \cite{etzioni_scanning_1967}. However, even a combination of these methods does not
precisely suit all classes of wicked problems. Is a problem tame? Use the \ac{classical-method}. Is it wicked, but
strategically under control and relatively non-controversial? Use \ac{incrementalism}. Is it wicked, and lacking a
strategic response or highly controversial? What then?

@@

Easy answers, or any solutions that artificially tame the problem, will not bring the matter under control. While by
definition there is no step that will truly solve the problem, easy answers deliberately emphasize one value to the
exclusion of others, leaving chaos in those neglected areas. As a result of ignoring root issues feeding the problem,
such solutions are even more likely to produce unintended consequences \cite{commission_tackling_2018}.

@@

Making sure every relevant group is included is important for two reasons. First, because information is ``uncertain and
diffuse,'' generating an accurate problem statement requires diverse input \cite{rozenshtein_wicked_2018}. Second, for
\ac{incrementalism} to work on the irreducibly wicked roots of the problem, there needs to be some degree of consensus
on overall strategy. Consensus building is not easy among groups with differing values and priorities, but it is
impossible without each group being represented.

% Prof. Engelsma
% - A big part of the "solution" is getting people together
% - My thesis seems aimed at tech people to get them to understand and judge threats
% - How can I take this accessible and bridge the gap for policymakers?

@@

The failures of the \ac{classical-method} and \ac{incrementalism} teach that we must be able to think strategically
while respecting the non-linear nature of \acp{wicked-prob}. One way to use the high level, holistic view of the
classical approach while intertwining the problem definition and analysis steps as in the incrementalist approach. In
practice, this means that the collective understanding of the problem and potential solutions must co-evolve. As Rittel
puts it, ``The systems-approach `of the first generation' [i.e., classical analysis] is inadequate for dealing with
wicked-problems. Approaches of the `second generation' should be based on a model of planning as an argumentative
process in the course of which an image of the problem and of the solution emerges gradually among the participants, as
a product of incessant judgment, subjected to critical argument'' \cite{rittel_dilemmas_1973}.

@@

This follows partially from the previous suggestion---problem definition and analysis are combined precisely because the
problem definition depends on the nature of proposed solutions. But this advice deserves emphasis for another reason:
consensus is easier to build around concrete proposals. While debate rages in the abstract, groups are bound to disagree
due to their conflicting values and priorities, but Lindblom writes encouragingly about ``the ease with which individuals
of different ideologies often can agree on concrete policy'' in an example about congressional compromise; he goes on to
say, ``Labor mediators report a similar phenomenon: the contestants cannot agree on criteria for settling their disputes
but can agree on specific proposals. Similarly, when one administrator's objective turns out to be another's means, they
often can agree on policy'' \cite{lindblom_muddling_1959}.

% So should I follow my own advice and focus on
%   group_2019 (argues for DAR) and
%   savage_lawful_2018 (a concrete DAR proposal)? -- and related, \cite{servan_schreiber_jje_2020}

@@

The above discussion describes the failure of well-established methods of policymaking at addressing \acp{wicked-prob}
and properties a better method would have. Here, I propose a modified \acs{OODAloop} as an alternative policymaking
model.

The \acl{OODAloop} was a developed by Air Force Colonel John Boyd as a description of a successful strategy for
countering opponents in real-time combat \cite{angerman_2004}. Boyd created it in the context of military strategy, but
its ideas have penetrated many other sectors, including cybersecurity, where it serves as a model for structuring
incident response \cite{schneier_future_2014}. There is an emphasis on fast cycle times and ``getting inside'' your
opponent's loop as a means of overcoming raw power with speed and agility. It is illustrated in \myfig{fig-ooda-loop}.

@@

The \ac{OODAloop} serves as a good basis for contending with \acp{wicked-prob} because it has the correct fundamental
structure. It maps very well to the \ac{classical-method} steps of problem definition (orientation), ideation and
analysis (orientation), decision, and action, but it emphasizes the \ac{incrementalism}'s iteration and feedback.
Additionally, while policymaking doesn't share the pace of real-time combat, in other ways the metaphor holds up. One is
facing an unpredictable opponent in a dynamic environment where every action counts. We cannot overpower
\acp{wicked-prob} through force of reason, but perhaps we can through speed and agility.

The one modification the \ac{OODAloop} needs is unification of the problem definition and analysis steps, as one depends
on the other. This is Rittel's ``argumentative process in the course of which an image of the problem and of the
solution emerges gradually'' \cite{rittel_dilemmas_1973}, illustrated in \myfig{fig-policy-ooda-loop} as a cycle of
``collaborative debate'' between observation and orientation. The modified \ac{OODAloop} incorporates each of the
lessons from \mysec{sec-policy-lessons}.

@@

Traditional policymaking typically follows either a rational approach rooted in the scientific method, aiming to be
thorough and complete, or an intuitional approach rooted in evolutionary trial and error, humbly aiming only to take
steps in the right direction. Both approaches have their strengths, but neither is suited to the complex and
controversial nature of so-called ``\acp{wicked-prob}.'' The strategy proposed here relies on the goodwill participation
of all parties, who engage in ongoing research and debate that rejects easy answers, encourages flexible risk-based
solutions, and crystallizes discussion around specific proposals. The result from a single iteration is an imperfect
step forward based on implementing a particular proposal. The result from many iterations is the breakdown of the
problem into subproblems---some tame and inevitably some still wicked. The tame subproblems can be addressed with the
\ac{classical-method}, and the wicked subproblems, by now restrained under a sound and agreeable strategy, can be
addressed with \ac{incrementalism}.

@@

This process is demonstrated in \myfig{fig-policy-ooda-process}. An out of control problem (the large storm) is being
confronted. Several iterations of the \ac{OODAloop} gradually bring it down to size, spinning out tame problems (puzzle
pieces) addressed by classical analysis and wicked but restrained problems (small storms) controlled by
\ac{incrementalism}.

@@

It is important to note that the same \acp{wicked-prob} is never solved twice, since each one is by definition, unique.
However, changes in technology, culture, and current affairs may render the current strategy insufficient and require
the process to start again. \Acp{wicked-prob} are like history: they don't repeat themselves, but they rhyme. Due to
changing circumstances, solutions past will not work in the future, but lessons learned from the past provide useful
experience.

In the context of \ac{encryption} and \ac{EA}, this means that the outcome of \ac{the-cw1} doesn't hold precedential
power over the debate in the second. It means that re-raising questions about the role of technology is playing society
is legitimate. To follow this proposed strategy would mean that hard-line rhetoric doesn't help \cite{ruiz_there_2018}
\cite{geller_2019}, though sincere appeals do \cite{abelson_2015} \cite{intl_2020} \cite{rozenshtein_2019}. Even more
helpful are collaborative efforts involving government, technical, and civil liberties representatives
\cite{committee_decrypting_2018} \cite{group_2019}. Specific proposals around which groups can center discussion,
whether they specify a particular form of \ac{EA} or perhaps especially if they don't, are also a necessary component
\cite{kerr_encryption_2017} \cite{wright_crypto_2018} \cite{phan_key_2017}.

>>>>

Before detailing the specific arguments used in the \ac{EA} debate, it is worth achieving a better understanding of
\acp{wicked-prob} in general. What are they, and why do the problems surrounding encryption, privacy, and \ac{EA} fall
into this category? What approaches to tackling \acp{wicked-prob} succeed and fail? How can we strategically confront
them and make real progress? This chapter seeks to answer these questions.

@@

\Acp{wicked-prob} were previously introduced as pernicious and tricky issues that resist straightforward solutions.
This section analyzes the nature of wicked problems and approaches to tackling them.

@@

Rittel's categorization of \acp{wicked-prob} grew out of frustration with their resistance to traditional problem
solving methods. Since the Enlightenment, society has applied the scientific method to problems of every kind; the
sweeping application of scientific analysis has delivered reliable clean water, improved crop yields, shaped government
structures, and bestowed material wealth previously unimaginable. With these material problems largely solved in the
twentieth century, believers in the power of reason thought this progress would continue in the realm of public
planning. Policymaking would function by setting goals, identifying problems, evaluating alternatives, implementing
solutions, and analyzing outcomes in order to correct errors. Functioning as a continuous process, this approach was
primed to revolutionize governing the same way it did industry, agriculture, and economics---until it didn't. In the
context of what he describes as an ``anti-professional movement,'' Rittel explains how the scientific method has failed:

@@

When applied to social problems, the prescribed method---here defined as setting goals, identifying problems, evaluating
alternatives, implementing solutions, and analyzing outcomes---fails at every step. In the U.S., a nation composed of
many varying cultures and subcultures and politically dominated by two mutually hostile parties, agreeing on
goals is a challenge in itself. When goals are set, we often discover that we are contending with \acp{wicked-prob} that
defy the process at each remaining step. Rittel provides a list of ten characteristics of problems in this category
\cite{rittel_dilemmas_1973}:

@@

In summary, (1) the \acp{wicked-prob} are impossible to definitively identify, (2) potential solutions are impossible to
definitively evaluate, (3) every action or inaction has permanent effects, and (4) solutions are importable from one
problem to another. The policy version of the scientific method cannot be used under these conditions. The ``wicked''
problem's counterpart is the ``tame'' problem. ``Tame'' does not imply easy; it only means that the scientific approach
will be effective.

Several of \acp{wicked-prob}' characteristics are results of the fact that we have no accurate predictive model of the
world and human behavior. This is self-evident. The more important insight is one step removed: precisely because there
is an inexhaustible set of potential solutions, the problem definition---that is, the set of information required to
produce a solution---is not self-contained. Each proposed solution demands that new research and context be fed back
into the problem definition. This feedback from proposal to definition violates the linearity of the traditional
approach. One cannot reason from problem statement to proposals to a solution; instead, one is forced to constantly
refine the problem statement based on the content of proposals themselves.

A 2018 study by the Australian government uses climate change as an example of a \ac{wicked-prob}
\cite{commission_tackling_2018}. Say the problem is initially formulated as long-term change to the environment caused
by the effects of accumulating greenhouse gasses. Responses typically take one of three broad forms
\cite{commission_tackling_2018}. First, profligate behavior in consumeristic societies must be reigned in at a local and
personal level. Second, global inter-governmental coordination is the only solution, as individual changes will have no
impact. Third, the situation is overblown by idealists and power-mongers, and technological progress and adaptive
markets will handle any negative effects that come to pass. Which response is correct? It is impossible to tell by
evaluating the problem statement. How much do individual choices contribute to greenhouse gas emissions? How effective
are international accords? How costly will the changes be, and how capable is technology to respond? Evaluating the
validity of each proposal requires updates to the problem statement itself.

@@

\Acp{wicked-prob} always consist of subproblems, which may be tame or wicked themselves. For example, climate change has
many subproblems. Developing clean energy technology is a tame subproblem; implementing any solution in a way that
doesn't leave vast swaths of people behind is a wicked subproblem. \Acp{wicked-prob} quickly grow in complexity with the
number of subproblems that comprise them. Truly imposing \acp{wicked-prob} are composed of a tangle of contributing
factors, and despite the presence of tame elements, they pose a large number of difficult challenges.

@@

\begin{enumerate}
  \item There is no formulation that encapsulates the problem of encryption's interplay with privacy, security,
        safety, and trust.
  \item Balancing each value in the face of constantly evolving technology is a never-ending cycle.
  \item \ac{EA} or other proposals cannot definitively solve the problem.
  \item \ac{EA} or other proposals cannot be objectively tested.
  \item Every policy implementation has irreversible effects.
  \item There is an inexhaustible set of potential solutions to the problem.
  \item Solutions from other domains do not apply directly to this problem.
  \item The need for \ac{EA} or some alternative is a symptom of differing values, rapid technological change, and
        criminal behavior.
  \item The problem can be framed many ways: as insufficient investigatory access to digital data (presupposing \ac{EA}
        as the solution), outdated cyberlaw (presupposing legal hacking or compelled password disclosure as solutions),
        and more.
  \item The decisions of regulators and technologists have real impacts in the world today.
\end{enumerate}

Cybersecurity law scholar Alan Rozenshtein argues at length for treating \ac{encryption} and \ac{EA} as a
\ac{wicked-prob} \cite{rozenshtein_wicked_2018}. He divides the root issues into three categories. First, there is
disagreement on what the goals should be (and even basic premises). Sides do not agree on whether things are ``going
dark'' or we're in the ``golden age for surveillance'' or how much to value competing notions of security. Second,
information is ``uncertain and diffuse.'' There are insufficient statistics on encryption's effect on investigations.
Though the consensus is that we are not presently capable of acceptably secure \ac{EA}, that is not certain to always be
the case, especially as the area is under-researched. Third, that the problem cannot be definitively solved. Evolving
values and technology mean that the subject is always up for renegotiation.

\Ac{encryption} technology is particularly sensitive to the irreversible effects of policy implementation. Technology
deployments have long tails, and attackers have the ability to record and store data for later analysis. In one
investigation, Australian authorities cracked a cold case based on evidence acquired from a mobile device that they
cracked five years after it was seized \cite{evans_arrests_2020}. Lawful authorities wouldn't be the only ones to
benefit from a policy miscalculation that could negatively impact security for a very long time.

@@

Embracing reality is the first step to dealing with it \cite{baker_2019}. We have by now embraced the reality of
\acp{wicked-prob}. The next two sections investigate strategies for dealing with it.

@@

The \ac{classical-method} \cite{feeley_judicial_2000} (also known as ``the modern-classical model of planning''
\cite{rittel_dilemmas_1973}, ``the rational-comprehensive method'' \cite{lindblom_muddling_1959}, ``traditional policy
analysis'' \cite{rozenshtein_wicked_2018}, or ``linear thinking'' \cite{commission_tackling_2018}) has already been
introduced. It is the reason-based method that functions by setting goals, identifying problems, evaluating
alternatives, implementing solutions, and analyzing outcomes in order to correct errors. \myfig{fig-classical-method}
illustrates this approach in the context of encryption and \ac{EA}. It has a purely linear flow from problem to solution
except for the ``refinement'' step, in which the method analyzes outcomes and corrects errors. However, it is important
to note that refinement is always in-kind---it represents a reinforcement, as opposed to a reassessment, of the chosen
solution.

@@

The shortcomings of the \ac{classical-method} were discussed in \mysec{wicked-characteristics}. It fails due to
disagreement over goals, the dependence of the problem definition on the potential solutions generated (violating the
linear flow), the inability to evaluate alternatives, and the lack of a definitive stopping rule.

@@

Lindblom's process of ``\ac{muddling-through}'' operates by taking successive steps chosen through comparative analysis.
The alternatives selected for comparison are defined relative to the status quo and are close enough to one another to
be able to be analyzed on the margin. This is done first out of practical necessity, due to our inability to rationally
predict policy outcomes, and second out of political realism, as non-incremental changes are usually politically
impossible to impose in a democratic system \cite{lindblom_muddling_1959}. The formality of the process varies;
policymakers may use this method consciously, with considerable comparative analysis, or unconsciously, led by
intuition. \myfig{fig-muddling-through} illustrates this method.

@@

Despite its strengths, \ac{incrementalism} also has weaknesses in its ability to address \acp{wicked-prob}. Its main
weakness is the absence of high level strategic analysis. Incrementalism is incapable of drastic change, which is
sometimes necessary. By Lindblom's admission, it lacks a safeguard for consideration of all relevant values and may
``overlook excellent policies for no other reason than that they are not suggested by the chain of successive policy
steps leading up to the present'' \cite{lindblom_muddling_1959}. Analysis based only on one present ``status quo'' to
the next can result in messy policy that none are particularly fond of---``As Lindblom's sobriquet suggests, it often
[leads] to a considerable muddle'' \cite{feeley_judicial_2000}. Application of the \acl{CFAA} is one such muddle
\cite{wolff_computer_2016}.

@@

The \ac{classical-method} and \ac{incrementalism} are not the only styles of policymaking. They represent perhaps two
extremes on a spectrum of rational planning and intuitive ``muddling.'' Both have strengths and weaknesses. One may
intuit that a reasonable strategy is the selective use of both approaches according to the situation, and indeed this
has been formally suggested \cite{etzioni_scanning_1967}. However, even a combination of these methods does not suit all
classes of wicked problems. Is a problem tame? Use the \ac{classical-method}. Is it wicked, but strategically under
control and relatively non-controversial? Use \ac{incrementalism}. What about when it is it wicked, lacks a strategic
response, and is highly controversial?

@@

Easy answers, or any solutions that artificially tame the problem, will not bring the matter under control. While by
definition there is no step that will truly solve the problem, easy answers deliberately emphasize one value to the
exclusion of others. This leads to chaos in those neglected areas. Solutions based on easy answers produce unintended
consequences because they ignore the root issues \cite{commission_tackling_2018}.

@@

Including every relevant group is important for two reasons. First, because information is ``uncertain and diffuse,''
generating an accurate problem statement requires diverse input \cite{rozenshtein_wicked_2018}. Second, for
\ac{incrementalism} to work on the irreducibly wicked roots of the problem, there needs to be some degree of consensus
on overall strategy. Consensus building is not easy among groups with differing values and priorities, but it is
impossible without each group being represented.

% TODO:
% Prof. Engelsma
% - A big part of the "solution" is getting people together
% - My thesis seems aimed at tech people to get them to understand and judge threats
% - How can I take this accessible and bridge the gap for policymakers?

@@

The failures of the \ac{classical-method} and \ac{incrementalism} reveal that we must be able to think strategically
while respecting the non-linear nature of \acp{wicked-prob}. One way to use the high level, holistic view of the
classical approach while intertwining the problem definition and analysis steps as in the incrementalist approach. In
practice, this means that the collective understanding of the problem and potential solutions must co-evolve. As Rittel
puts it, ``The systems-approach `of the first generation' [i.e., classical analysis] is inadequate for dealing with
wicked-problems. Approaches of the `second generation' should be based on a model of planning as an argumentative
process in the course of which an image of the problem and of the solution emerges gradually among the participants, as
a product of incessant judgment, subjected to critical argument'' \cite{rittel_dilemmas_1973}.

@@

Focusing on concrete proposals follows partially from the previous suggestion---problem definition and analysis are
combined precisely because the problem definition depends on the nature of proposed solutions. But this advice deserves
emphasis for another reason: consensus is easier to build around concrete proposals. While debate rages in the abstract,
groups are bound to disagree due to their conflicting values and priorities. However, Lindblom writes encouragingly
about ``the ease with which individuals of different ideologies often can agree on concrete policy'' in an example
about congressional compromise. He goes on to say, ``Labor mediators report a similar phenomenon: the contestants cannot
agree on criteria for settling their disputes but can agree on specific proposals. Similarly, when one administrator's
objective turns out to be another's means, they often can agree on policy'' \cite{lindblom_muddling_1959}.

@@

The discussion above describes the inability of well-established methods of policymaking to address \acp{wicked-prob}
and properties a better method would have. Here, I propose a modified \acs{OODAloop} as an alternative policymaking
model.

The \acl{OODAloop} was developed by Air Force Colonel John Boyd as a description of a successful strategy for countering
opponents in real-time combat \cite{angerman_2004}. Boyd created it in the context of military strategy, but its ideas
have penetrated many other sectors, including cybersecurity, where it serves as a model for structuring incident
response \cite{schneier_future_2014}. There is an emphasis on fast cycle times and ``getting inside'' your opponent's
loop as a means of overcoming raw power with speed and agility. It is illustrated in \myfig{fig-ooda-loop}.

@@

The \ac{OODAloop} serves well as a model for contending with \acp{wicked-prob} because it has the correct fundamental
structure. It maps very well to the \ac{classical-method} steps of problem definition (orientation), ideation and
analysis (orientation), decision, and action; however, it also emphasizes \ac{incrementalism}'s iteration and feedback.
Additionally, while policymaking does not share the pace of real-time combat, this metaphor holds true in other ways.
One is facing an unpredictable opponent in a dynamic environment in which every action counts. We cannot overpower
\acp{wicked-prob} through force of reason, but perhaps we can through speed and agility.

The OODA Loop does need one modification in order to suit wicked problems. Because the problem definition step and
analysis step depend on one another, they should be unified. This is Rittel's ``argumentative process in the course of
which an image of the problem and of the solution emerges gradually'' \cite{rittel_dilemmas_1973}, illustrated in
\myfig{fig-policy-ooda-loop} as a cycle of ``collaborative debate'' between observation and orientation. The modified
\ac{OODAloop} incorporates each of the lessons from \mysec{sec-policy-lessons}.

@@

Traditional policymaking typically follows either a rational approach rooted in the scientific method, aiming to be
thorough and complete, or an intuitional approach rooted in evolutionary trial and error, humbly aiming only to take
steps in the right direction. Both approaches have strengths, but neither is suited to the complex and controversial
nature of so-called ``\acp{wicked-prob}.'' The strategy proposed here relies on the goodwill participation of all
parties, who engage in ongoing research and debate that rejects easy answers, encourages flexible risk-based solutions,
and crystallizes discussion around specific proposals. The result of a single iteration is an imperfect step forward
based on implementing a particular proposal. The result of many iterations is the breakdown of the problem into
subproblems---some tame and inevitably some still wicked. The tame subproblems can be addressed with the
\ac{classical-method}, and the wicked subproblems, by now restrained under a sound and agreeable strategy, can be
addressed with \ac{incrementalism}.

@@

This process is demonstrated in \myfig{fig-policy-ooda-process}. An out-of-control problem (the large storm) is
confronted in this diagram. Several iterations of the \ac{OODAloop} gradually diminish its size. This is achieved by
spinning out tame problems (puzzle pieces) addressed by classical analysis and wicked but restrained problems (small
storms) controlled by \ac{incrementalism}.

@@

It is important to note that the same \ac{wicked-prob} is never faced twice, since each one is by definition, unique.
However, changes in technology, culture, and current affairs may render the current strategy insufficient and require
the process to start again. \Acp{wicked-prob} are like history: they don't repeat themselves, but they rhyme. Due to
changing circumstances, past solutions cannot be used in the future, but lessons learned from the past can be.

In the context of \ac{encryption} and \ac{EA}, this means that the outcome of \ac{the-cw1} doesn't hold precedential
power over the debate in the second. It means that re-raising questions about the role of technology plays in society is
legitimate. To follow this proposed strategy would mean that hard-line rhetoric doesn't help \cite{ruiz_there_2018}
\cite{geller_2019} but sincere appeals do \cite{abelson_2015} \cite{intl_2020} \cite{rozenshtein_2019}. Even more
helpful are collaborative efforts involving government, technical, and civil liberties representatives
\cite{committee_decrypting_2018} \cite{group_2019}. Specific proposals around which groups can center discussion,
whether they specify a particular form of \ac{EA} or perhaps especially if they don't, are also a necessary component
\cite{kerr_encryption_2017} \cite{wright_crypto_2018} \cite{phan_key_2017}.
